Deriving Multi-Headed Projective Dependency Parses \\ from Link Grammar Parses

### 2014-05-27 doc/LinksToDAG.tex

\begin{abstract}
% Copied from the google doc
There has been recent theoretical work on multi-headed versions of dependency grammar(cite). Such formalisms can be more syntactically and semantically expressive. However, there aren't any corpora at present (check this) except for Buch-Kromann's work on Danish \cite{kromann2003danish}, so it is hard to investigate the benefit of such parsers or to work on making them faster or more accurate. To fill this gap, we observe that link grammar produces parses that are similar to multi-headed planar dependency parses except that the links are undirected. We use Integer Linear Programming to assign consistent directions to the links in a corpus of NNN parses produced by the Link Grammar Parser, which has broad-coverage hand-written grammars of English, Russian, and more. (what statistics does it use?). We find that such directions can indeed be consistently assigned in a way that yields valid multi-headed dependency parses. The resulting parses in English appear linguistically plausible, although they are not in general consistent with CoNLL-style parses of the same sentences; we discuss the differences. We also report ...
\end{abstract}

\section{Introduction}
\TODO{Intro}
Projective dependency parsing is the task of mapping a sentence to a directed acyclic tree. Link parsing in contrast produces a multiheaded planar graph with undirected edges, where every edge has a label describing the relationship between two constituents in a parse. In this paper we explore whether these relationships include dependencies, and whether the multiheadedness of the link grammar offers additional dependency relationships not found in other corpora.

@@

Using the sentences of a dependency corpus as comparison, we find that the link parses do not wholly subsume dependency parses, and that the undirected links match roughly two thirds of the arcs in the conll dependency corpus. \TODO{This is because...} 
Of the links that match a corresponding dependency arc we measure whether the direction we chose is the same as the original dependency data. 
Finally, we look to see that given a link label, whether the original conll label can be recovered.

@@

@@

The link grammar is based on a set of handwritten dictionaries. Instead of going through these dictionaries, we learned the grammar using an ILP. This approach also potentially allows us to analyze link grammar dictionaries other than English. \TODO{explore other link grammar dictionaries}

@@

\TODO{cite model minimization papers}

@@

@@

\begin{figure}[ht!]
  \centering
  \small
  \input{figure/conll_analysis_sentences.tex}
  \caption{\small Of the sentences used for this experiment section about \input{figure/conll_analysis_sentences_dropped}had at least one dropped word, and about \input{figure/conll_analysis_sentences_multiheaded}had at least one multiheaded word}
\end{figure}

@@

\TODO{A table here that lists each link grammar type together with its
  count, its allowed direction(s), its percentage of majority
  directions by token, and its most common corresponding CoNLL tags
  with percentages.  Readers could cut-and-paste this data to make a 
  cheap post-processor for LG parses.}

@@

\TODO{Show a bunch of English sentences with CoNLL vs.\@ link parses.
  Explain how they were selected and how the results are colorized etc.}

>>>>

\begin{abstract}
Under multi-headed dependency grammar, a parse is a DAG rather than a tree.  Such formalisms can be more syntactically and semantically expressive.  However, there are few multi-headed corpora at present, particularly for the projective case, so it is hard to investigate the benefit of such parsers or to work on making them faster or more accurate.  To fill this gap, we observe that link grammar produces {\em undirected} projective graphs.  
% OLD
% ... produces parses that are similar to multi-headed projective dependency parses, except that the links are undirected. 
We use Integer Linear Programming to assign consistent directions to the labeled links in a corpus of !!! \noteJE{fill in the number} parses produced by the Link Grammar Parser, which has broad-coverage hand-written grammars of English, Russian, and more. \noteJE{what statistics does it use?} We find that such directions can indeed be consistently assigned in a way that yields valid multi-headed dependency parses. The resulting parses in English appear reasonably linguistically plausible, although they are not in general consistent with CoNLL-style parses of the same sentences; we discuss the differences.  We also report on the convergence of our ILP method to a stable annotation scheme as the corpus size increases.
\end{abstract}

\section{Introduction}
\todo{Intro}
Projective dependency parsing is the task of mapping a sentence to a directed acyclic tree. Link parsing in contrast produces a multiheaded planar graph with undirected edges, where every edge has a label describing the relationship between two constituents in a parse. In this paper we explore whether these relationships include dependencies, and whether the multiheadedness of the link grammar offers additional dependency relationships not found in other corpora.

@@

Using the sentences of a dependency corpus as comparison, we find that the link parses do not wholly subsume dependency parses, and that the undirected links match roughly two thirds of the arcs in the conll dependency corpus. \todo{This is because...} 
Of the links that match a corresponding dependency arc we measure whether the direction we chose is the same as the original dependency data. 
Finally, we look to see that given a link label, whether the original conll label can be recovered.

@@

Dependency parsing is the task of mapping a sentence to a projective (not always projective?) directed acyclic tree. Link parsing in contrast produces a multiheaded planar graph with undirected edges, where every edge has a label describing the relationship between two constituents in a parse. In this paper we explore whether these relationships include dependencies. To determine the directional dependencies within the link edge labels we will use integer linear programming, encoding the problem in the Zimpl little language \cite{Koch2004}. It turns out that the link parses roughly only match half of the conll dependency corpus. However this is because \todo{}.

Some recent work has considered multi-headed ...

 by expressing the dependency of a word on multiple parents.

\cite{mcdonald-pereira-2006,buchkromann-2006,sagae-tsujii-2008,gomezrodriguez-nivre-2013}

 (check this) except for Buch-Kromann's work on Danish \cite{kromann-2003}

@@

The link grammar is based on a set of handwritten dictionaries. Instead of going through these dictionaries, we learned the grammar using an ILP. This approach also potentially allows us to analyze link grammar dictionaries other than English. \todo{explore other link grammar dictionaries}

@@

\todo{cite model minimization papers}

@@

\subsection{Slack Hierarchy}

We introduced slack on the directional dependency variable tokens such that link types were allowed to deviate from the majority up to (1\% ?) of the time before the ILP would assign both directional dependency variables to true. This addressed noise in the link parser's label assignments, while still allowing for the possibility that both directions could still be assigned.

\todo{} We also introduced slack on the link types such that link types with the same coarse grained label would try and align the same way as the majority in the group, where the preceding capital letters of the link type denote the coarse grained label, while the subscript letters denote further information. This slack places a prior on rare or never-before seen link types to be assigned in the same way as other similar link types.

This slack hierarchy gave the ILP the flexibility to handle noise and novel link types while still trying to learn the overall link grammar.

\begin{equation}
  count_{L, \textsc{left}} = \sum_{a \in \textsc{ARCS}_L} left_a
\end{equation}
\begin{equation}
  count_{L, \textsc{left}} = \textsc{M} \cdot \textsc{allowed}_{L,\textsc{left}} + \textsc{slack}_{L}
\end{equation}
Where $L \in \textsc{LABELS}$, and $\textsc{M}$ is the size of the corpus.
The objective is:
\begin{equation}
  \min \sum_{L \in \textsc{LABELS} \atop \textsc{DIR} \in \{left, right\}} (\textsc{allowed}_{L,\textsc{DIR}}) \atop + \textsc{COST}(L) * \sum_L \textsc{slack}_L
\end{equation}
\begin{equation}
  \textsc{COST}(L) = \frac{100}{(\textsc{count}_{L,\textsc{left}} + \textsc{count}_{L,\textsc{right}})}
\end{equation}

\begin{algorithm}
\caption{ILP Encoding}\label{ILP}
\begin{algorithmic}[1]
%\Procedure{MyProcedure}{}

\State $\text{min} \sum_{l \in labels} allowed_L(l) + allowed_R(l)$
\State $\text{subject to}$
\State $\text{Acyclicity}$
\State $\text{All nodes reachable from the root.}$
\end{algorithmic}
\end{algorithm}

@@

\begin{figure}[ht!]
  \centering
  \small
  \input{figure/conll_analysis_sentences.tex}
  \caption{\small Of the sentences used for this experiment section, about \input{figure/conll_analysis_sentences_dropped}
    had at least one dropped word, and about \input{figure/conll_analysis_sentences_multiheaded}
    had at least one multiheaded word}
\end{figure}

@@

\todo[inline]{A table here that lists each link grammar type together with its
  count, its allowed direction(s), its percentage of majority
  directions by token, and its most common corresponding CoNLL tags
  with percentages.  Readers could cut-and-paste this data to make a 
  cheap post-processor for LG parses.}

@@

\todo[inline]{Show a bunch of English sentences with CoNLL vs.\@ link parses.
  Explain how they were selected and how the results are colorized etc.}
